Language-Based Security for Low-Level MPC

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

Given these definitions, we can formulate an invariant
for circuit computation with respect to internal gates
as follows. It says that the output of any gate is
deterministic wrt inputs $S$, and conditionally
on $S$ the output $\macgv{\mesg{z}}$ remains
separable from corrupt views and both shares of
$\macgv{\mesg{z}}$. This last nuance is critical,
since those shares will in fact be revealed if
$\macgv{\mesg{z}}$ is decoded as the circuit output. 
\begin{lemma}[GMW Invariant]
  \label{lemma-gmwinvariant}
  Given:
  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
  Then all of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
  \begin{enumerate}
  \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
  \item $\condunix{\progtt(\prog)}{S}{M_C}$
  \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
  \end{enumerate}
\end{lemma}
To prove this, we can formulate and automatically prove gate-level
versions of the invariant. This serves as a proof tactic
that simplifies the proof of the GMW invariant. It establishes
that the $\eand$ gate output is deterministic conditional on
the inputs, and is separable from the output shares and
either parties' received messages. 
\begin{lemma}[And Gate Tactic]
  \label{lemma-gmwtactic}
  %Define:
  %$$
  %\begin{array}{c}
  %  \prog_{i} \defeq \\
  %  \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
  %  \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
  %\end{array}
  %$$
  Given:
  $$
  \begin{array}{c}
  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
  (\prog,\mesg{z})
  \end{array}
  $$
  Then all of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
  \begin{enumerate}
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
\end{lemma}
\begin{proof}
Verified automatically using techniques described in Section \ref{section-bruteforce}.  
\end{proof}

To properly integrate the local reasoning of Lemma \ref{lemma-gmwtactic} with
the global reasoning of Lemma \ref{lemma-gmwinvariant}, we can demonstrate
the following.
\begin{lemma}
  \label{lemma-conditioning}
  Give $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
  \begin{enumerate}
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_2}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_1}{\unix{V_2}}$, then $\condp{\pmf}{S}{\unix{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condsep{\pmf}{V_1}{V_2}{V_3}$, then $\condsep{\pmf}{S}{V_2}{V_3}$.
    %\item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
    %  and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
    %\item Given $\condsep{\pmf}{S_1 \cup S_2}{V_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
    %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
  \end{enumerate}
\end{lemma}

Then we can put the pieces together to prove the invariant, using automated tactics
for gate-level reasoning.  
\begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
  By induction on the length of $(\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}]))$.
  Encoding establishes the invariant in the basis. The most interesting inductive
  case is the $\andgmw$ gate. 
  \paragraph{Case $\andgmw$.} In this case we have:
  $$
  \begin{array}{c}
  (\varnothing,e) \redxs (\prog',\decodegmw(E[\andgmw(z,\mesg{x},\mesg{y})])) \redxs \\
    (\prog,\decodegmw(E[\mesg{z}]))
  \end{array}
  $$
  Letting $\iov(\prog) = S \cup M$, by definition and the induction hypothesis there
  exist $S^1,S^2 \subseteq S$ and $M^1,M^2 \subseteq M$
  such that:
  \begin{enumerate}[\hspace{5mm}(x.1)]
  \item $\conddetx{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}$
  \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
  %\item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
  \end{enumerate}
  and also:
  \begin{enumerate}[\hspace{5mm}(y.1)]
  \item $\conddetx{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}$
  \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
  %\item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
  \end{enumerate}
  and by Lemma \ref{lemma-gmwtactic} we have:
  \begin{enumerate}[\hspace{5mm}(z.1)]
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
  Given $x.1$ and $y.1$ and assumed uniqueness of random variables
  used in each gate we have $\conddetx{\progtt(\prog)}{S}{\{
    \macgv{\mesg{x}},\macgv{\mesg{y}}\}}$, so it follows from $z.1$
  and Lemma \ref{lemma-conditioning} (1) that
  $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$, and also by
  $z.3$ and Lemma \ref{lemma-conditioning} (3):
  $$\condsep
  {\progtt(\prog)}
  {S}
  {\{ \macgv{\mesg{z}} \}}
  {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$$
  Additionally Lemma \ref{lemma-conditioning} (2) gives $\condunix{\progtt(\prog)}{S}{\{ \elab{\mesg{z}}{\cid} \}}$,
  and $x.2$ and $y.2$ give $\condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C}$, $z.3$ together with
  uniqueness of gate identifiers implies:
  $$
  \condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C \cup \{  \elab{\mesg{z}}{\cid} \}}
  $$
  This implies the result.
\end{proof}

\begin{theorem}
  \label{theorem-gmw}
  If $e$ is a GMW circuit protocol in $\metaprot$ with $(\varnothing,e) \redxs (\prog,())$
  then $\prog$ satisfies noninterference modulo output. 
\end{theorem}

>>>>

Given these definitions, we can formulate an invariant for circuit
computation with respect to internal gates as follows. It says that
the output of any gate is deterministic given inputs $S$, and
conditionally on $S$ corrupt views are always in a uniform random
disrtibution (pure noise), while output $\macgv{\mesg{z}}$ remains
separable from corrupt views and both shares of
$\macgv{\mesg{z}}$. This last condition (incidentally missing from PSL
due to the reliance on more recent innovations in conditioning logic
\cite{li2023lilac}) is critical since those shares will in fact be
revealed if $\macgv{\mesg{z}}$ is decoded as the circuit output.
\begin{lemma}[GMW Invariant]
  \label{lemma-gmwinvariant}
  Given:
  $$ (\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}])) $$
  Then all of the following conditions hold for all $H$ and $C$ where $\iov(\prog) = S \cup M$:
  \begin{enumerate}
  \item $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$
  \item $\condunix{\progtt(\prog)}{S}{M_C}$
  \item $\condsep{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}{\{ \elab{\mesg{z}}{1}, \elab{\mesg{z}}{2} \})}$
  \end{enumerate}
\end{lemma}
To prove this, we can formulate and automatically prove local,
gate-level versions of the invariant. This serves as a proof tactic
that simplifies the proof of the GMW invariant. 
\begin{lemma}[And Gate Tactic]
  \label{lemma-gmwtactic}
  %Define:
  %$$
  %\begin{array}{c}
  %  \prog_{i} \defeq \\
  %  \eassign{\mesg{x}}{1}{\flip{x}}{1}; \eassign{\mesg{x}}{2}{\flip{x}}{2}; \\
  %  \eassign{\mesg{y}}{1}{\flip{y}}{1}; \eassign{\mesg{y}}{2}{\flip{y}}{2} 
  %\end{array}
  %$$
  Given:
  $$
  \begin{array}{c}
  (\varnothing,\andgmw(z,\mesg{x},\mesg{y}) \redxs %\\
  (\prog,\mesg{z})
  \end{array}
  $$
  Then all of the following conditions hold for both $\cid \in \{ 1,2 \}$ where $\iov(\prog) = M$:
  \begin{enumerate}
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
\end{lemma}
\begin{proof}
Verified automatically using techniques described in Section \ref{section-bruteforce}.  
\end{proof}

To properly integrate the local reasoning of Lemma \ref{lemma-gmwtactic} with
the global reasoning of Lemma \ref{lemma-gmwinvariant}, we can demonstrate
the following properties of conditioning-based reasoning. While a frame rule
was developed in \cite{li2023lilac}, these properties are distinct and
useful for reasoning about MPC. We omit proofs for brevity. 
\begin{lemma}
  \label{lemma-conditioning}
  Give $\pmf$, each of the following hold for $S,V_1,V_2,V_3 \in \dom(\pmf)$:
  \begin{enumerate}
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_2}{\detx{V_2}}$, then $\condp{\pmf}{S}{\detx{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condp{\pmf}{V_1}{\unix{V_2}}$, then $\condp{\pmf}{S}{\unix{V_2}}$.
    \item Given $\condp{\pmf}{S}{\detx{V_1}}$ and
      $\condsep{\pmf}{V_1}{V_2}{V_3}$, then $\condsep{\pmf}{S}{V_2}{V_3}$.
    %\item Given $\condsep{\pmf}{X_1}{X_2}{X_3}$ and $\condp{\pmf}{X_1}{\detx{X_2}}$
    %  and $\condp{\pmf}{X_2}{\detx{X_4}}$, then $\condsep{\pmf}{X_1}{X_4}{X_3}$.
    %\item Given $\condsep{\pmf}{S_1 \cup S_2}{V_3}{X_4}$ and $\condp{\pmf}{X_1 \cup X_2}{\detx{X_3}}$,
    %  then $\condsep{\pmf}{X_1}{X_2 \cup X_3}{X_4}$.
  \end{enumerate}
\end{lemma}

Then we can put the pieces together to prove the invariant, using automated tactics
for gate-level reasoning. We sketch some elements of this proof to focus on
novel aspects of our technique. 
\begin{proof}[Proof of Lemma \ref{lemma-gmwinvariant}]
  By induction on the length of $(\varnothing,e) \redxs (\prog,\decodegmw(E[\mesg{z}]))$.
  Encoding establishes the invariant in the basis. The most interesting inductive
  case is the $\andgmw$ gate. 
  \paragraph{Case $\andgmw$.} In this case we have:
  $$
  \begin{array}{c}
  (\varnothing,e) \redxs (\prog',\decodegmw(E[\andgmw(z,\mesg{x},\mesg{y})])) \redxs \\
    (\prog,\decodegmw(E[\mesg{z}]))
  \end{array}
  $$
  Letting $\iov(\prog) = S \cup M$, by definition and the induction hypothesis there
  exist $S^1,S^2 \subseteq S$ and $M^1,M^2 \subseteq M$
  such that:
  \begin{enumerate}[\hspace{5mm}(x.1)]
  \item $\conddetx{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}$
  \item $\condunix{\progtt(\prog)}{S^1}{M^1_C}$
  %\item $\condsep{\progtt(\prog)}{S^1}{\{\macgv{\mesg{x}}\}}{\{ \elab{\mesg{x}}{1}, \elab{\mesg{x}}{2} \})}$
  \end{enumerate}
  and also:
  \begin{enumerate}[\hspace{5mm}(y.1)]
  \item $\conddetx{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}$
  \item $\condunix{\progtt(\prog)}{S^2}{M^2_C}$
  %\item $\condsep{\progtt(\prog)}{S^2}{\{\macgv{\mesg{y}}\}}{\{ \elab{\mesg{y}}{1}, \elab{\mesg{y}}{2} \})}$
  \end{enumerate}
  and by Lemma \ref{lemma-gmwtactic} we have:
  \begin{enumerate}[\hspace{5mm}(z.1)]
  \item
    $\conddetx{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \macgv{\mesg{z}} \}}$
  \item $\condunix{\progtt(\prog)}{\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}{\{ \elab{\mesg{z}}{\cid} \}}$
  \item $\condsep
    {\progtt(\prog)}
    {\{ \macgv{\mesg{x}},\macgv{\mesg{y}} \}}
    {\{ \macgv{\mesg{z}} \}}
    {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$
  \end{enumerate}
  Given $x.1$ and $y.1$ and assumed uniqueness of random variables
  used in each gate we have $\conddetx{\progtt(\prog)}{S}{\{
    \macgv{\mesg{x}},\macgv{\mesg{y}}\}}$, so it follows from $z.1$
  and Lemma \ref{lemma-conditioning} (1) that
  $\conddetx{\progtt(\prog)}{S}{\{\macgv{\mesg{z}}\}}$, and also by
  $z.3$ and Lemma \ref{lemma-conditioning} (3):
  $$\condsep
  {\progtt(\prog)}
  {S}
  {\{ \macgv{\mesg{z}} \}}
  {\{ \elab{\mesg{z}}{1},\elab{\mesg{z}}{2} \}}$$
  Additionally Lemma \ref{lemma-conditioning} (2) gives $\condunix{\progtt(\prog)}{S}{\{ \elab{\mesg{z}}{\cid} \}}$,
  and $x.2$ and $y.2$ give $\condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C}$, $z.3$ together with
  uniqueness of gate identifiers implies:
  $$
  \condunix{\progtt(\prog)}{S}{(M^1 \cup M^2)_C \cup \{  \elab{\mesg{z}}{\cid} \}}
  $$
  This implies the result.
\end{proof}
The preceding Lemma, together with some additional observations about decoding, establish correctness
of arbitrary circuits. 
\begin{theorem}
  \label{theorem-gmw}
  If $e$ is a GMW circuit protocol in $\metaprot$ with $(\varnothing,e) \redxs (\prog,())$
  then $\prog$ satisfies noninterference modulo output. 
\end{theorem}
